---
title: "scHiCARdownstream Vignette: scHiCAR downstream analysis via ArchR"
author: "Jianhong Ou"
date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('scHiCARdownstream')`"
abstract: >
  scHiCARdownstream is a R package for scHiCAR downstream analysis via ArchR.
vignette: >
  %\VignetteIndexEntry{scHiCARdownstream Vignette: scHiCAR downstream analysis via ArchR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  html_document:
    theme: simplex
    toc: true
    toc_float: true
    toc_depth: 4
    fig_caption: true
---

```{r, echo=FALSE, results="hide", warning=FALSE}
knitr::opts_chunk$set(
    warning = FALSE, message = FALSE,
    fig.width = 12, fig.height = 4,
    dpi = 72
)
```
## Introduction




## Installation

```{r, eval=FALSE}
if (!require("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}
BiocManager::install("jianhong/scHiCARdownstream")
```

## Quick start

There are five main steps to use `scHiCARdownstream` for downstream analysis of scHiCAR data:

Step 1. Load the required library.
```{r loadlibrary,warning = FALSE,message=FALSE}
library(scHiCARdownstream)
library(ArchR)
library(BSgenome.Mmusculus.UCSC.mm10)
```

Step 2. Read RNA-DNA barcode map. 

```{r bmap}
barcodeMap <- read.csv(system.file('extdata', 'barcode.map.csv', package='scHiCARdownstream'))
bmap <- barcodeMap$`DNA`
names(bmap) <- barcodeMap$`RNA`
```

Step 3. Create ArchR project


```{r createProj}
extdata <- system.file('extdata', package = 'scHiCARdownstream')
rnaFiles <- dir(file.path(extdata, 'RNA'), 'matrix.mtx.gz',
                full.names=TRUE, recursive=TRUE)
atacFiles <- dir(file.path(extdata, 'ATAC'), 'ATAC.tsv.gz$',
                 full.names = TRUE)
names(rnaFiles) <- names(atacFiles) <- 'dpa7'
addArchRGenome("mm10")
proj <- createArchRproj(
    rnaFiles, atacFiles,
    bmap, barcodesFun=function(x){
      do.call(rbind, strsplit(colnames(x), '_'))[, 4]
    },
    outputDirectory = "ArchR_proj",
    annotationGenes = NULL,
    verbose=TRUE
    )
```
Step 4. Process the data

```{r processArchR}
proj <- processArchR(proj, skip_rmDup=TRUE) ## data are not good for rm doublets.
plotEmbedding(proj, 
              name = "Clusters_Combined",
              embedding = "UMAP_Combined",
              size = 1, labelAsFactors=F, labelMeans=F)
```
```{r callpeak, eval=FALSE}
# and call peaks, make sure macs2 available or provide the path via parameter `pathToMacs2`.
proj <- callPeaks(proj)
```

Step 5. prepare files for [FastHigashi](https://github.com/ma-compbio/Higashi/wiki/Input-Files)

```{r prepareHigashiFiles}
pairs_files <- dir(file.path(extdata, 'Pairs'), 'pairs.gz', full.names = TRUE)
names(pairs_files) <- 'dpa7'
prepareHigashiFiles(proj, pairs_files)
```
Step 6. run FastHigashi (ask Higashi installed)

This step as Higashi and FastHigashi installed. The parameter `use_conda` is still in test.
```{r runHigashi, eval=FALSE}
runFastHigashi() # runFastHigashi(use_conda='higashi_env')
```

Step 7. prepare files for [scDeepLUCIA](https://github.com/kaistcbfg/scDeepLUCIA)

Generate the bigwig files for ATAC reads for each sample
```{r createBW}
ArchR::getGroupBW(proj, groupBy = "Sample",
                  normMethod = "nFrags",
                  tileSize = 25, # tileSize 25 is required by scDeepLUCIA
                  maxCells = 1000)
```
Run scripts to create the Epi array files. Note the script ask `pyBigWig` and `numpy` installed.
```{r eval=FALSE}
# Retrieve path to the script within your package
script_path <- system.file("extdata", "scripts", "prepare_scDeepLUCIA_epi.py",
                           package = "scHiCARdownstream")
# rewrite the sys.argv for argparse.ArgumentParser
sys <- reticulate::import("sys")
sys$argv <- list(
    script_path,
    "--bw_roots", file.path(ArchR::getOutputDirectory(proj), 'GroupBigWigs', 'Sample'),
    "--marker_name", 'r2_030M',
    "--resolution", 5000,
    "--outdir", "scDeepLUCIA_inputs/sliced_epi_array/mm10",
    "--nan_to", 0.0
  )
# Run the Python script
reticulate::py_run_file(script_path)
```

Create one-hot encoded genomic feature npy file.

```{r eval=FALSE}
createOnehotGenomicNpy(proj, resolution=5000)
```

Create 5kb-resolution contact matrix npy file.

```{r eval=FALSE}
createContactMatrix(proj, pairs_files)
```


## SessionInfo

```{r pressure, echo=FALSE}
sessionInfo()
```

